---
title: "p8105_hw1_yg3099"
author: "daisy_gui"
date: "2025-09-21"
output: github_document
---

## Problem 1

```{r setup}
library("moderndive")
library(ggplot2)
```

```{r load_data}
data("early_january_weather") #load data
```

This dataset include the following variables: `r colnames(early_january_weather)`. 

One of the important variables is 'temp', which records the temperature at the time of record.

Another important variable is 'time_hour', which records the time that the weather information are measured and recorded. 

The dataset include `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. 

The mean temperature among the collected data is `r mean(early_january_weather$temp)`. 

```{r}
# create and print scatterplot
temp_time_plot = ggplot(
  early_january_weather, 
  aes(x = time_hour, y = temp, color = humid)
  ) + 
  geom_point()

print(temp_time_plot)

# save plot to directory
ggsave(filename = "temp_time_scatterplot.png", plot = temp_time_plot)

```

